---
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: inline
---
Red Wine Quality  Exploration by Hani Jubail
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
library(memisc)


```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
getwd()
df<- read.csv("C:/Users/PC/Desktop/R/wineQualityReds.csv")
```

> **introduction**:  The dataset were created, using red wine samples.
  The inputs include objective tests (e.g. PH values) and the output is based on sensory data
  (median of at least 3 evaluations made by wine experts). Each expert graded the wine quality 
  between 0 (very bad) and 10 (very excellent).

# Univariate Plots Section


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_1}

head(df)
```
> The quality must be converted into factors because it is a categorical variable .

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_as_factors}

df$quality <- as.factor(df$quality)

head(df)
```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_2}

ggplot(aes(x = quality),
       data = df) +
  geom_histogram(stat="count")

```

> Most of the red wine from the sample are of level 5 and 6 quality, with fewer samples of lesser-quality and higher levels at 7 or 8.



```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_3}

ggplot(aes(x = alcohol),
       data = df)+
  geom_histogram(binwidth = 0.3)

```
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_4}

summary(df$alcohol)

```
> The mean alcohol content of the wine is at 10% and it's right-skewed  distribution. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_5}

ggplot(aes(x =pH),
       data = df) +
  geom_histogram()

```
```{r echo=FALSE,  message=FALSE, warning=FALSE,Univariate_Plots_6}

summary(df$pH)

```
> on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_7}

ggplot(aes(x = density),
       data = df) +
  geom_histogram()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_8}

summary(df$density)

```
> the density of water is close to that of water depending on the percent alcohol and sugar content and we can see from the plot and the mean is 0.9967 .

```{r echo=FALSE,  message=FALSE, warning=FALSE,Univariate_Plots_9}

ggplot(aes(x =sulphates),
       data = df) +
  geom_histogram()

```
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_10}

summary(df$sulphates)

```
> Sulphates values above 1.5 are outliers.

```{r echo=FALSE,  message=FALSE, warning=FALSE, Univariate_Plots_11}

ggplot(aes(x =chlorides),
       data = df) +
  geom_histogram()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_12}

summary(df$chlorides)

```

> chlorides is the amount of salt in the wine and chlorides more than 0.6 are outliers.


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_13}

ggplot(aes(x =residual.sugar),
       data = df) +
  geom_histogram()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_14}

summary(df$residual.sugar)

```

> the amount of sugar remaining after fermentation stops is around 2 grams/ 1 litre.


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_15}

ggplot(aes(x =fixed.acidity),
       data = df) +
  geom_histogram()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_16}
summary(df$fixed.acidity)


```



```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_17}

ggplot(aes(x =volatile.acidity),
       data = df) +
  geom_histogram()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_18}

summary(df$volatile.acidity)

```

# Univariate Analysis

### What is the structure of your dataset?

Number of Instances:1599,Number of Attributes: 12.

1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily).

   2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste.

   3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines.

   4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet.

   5 - chlorides: the amount of salt in the wine.

   6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine.

   7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine.

   8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content.

   9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale.

   10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant.

   11 - alcohol: the percent alcohol content of the wine.

   12 - quality (score between 0 and 10).
   
   #https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt
   
### What is/are the main feature(s) of interest in your dataset?
The main features in the data set are quality , alcohol and residual sugar.
I'd like to determine which features are best for predicting the the quality of the wine.

### What other features in the dataset do you think will help support your \
density, pH, aciditi"fixed or volatile"and chlorides can increase or decrease  the quality of the wine .

### Did you create any new variables from existing variables in the dataset?\
not yet but i might group some variables in the Bivariate section.

### Of the features you investigated, were there any unusual distributions? \
there is no unusal disributions, most of them are normal or  right-skewed distributions with some outliers.

# Bivariate Plots Section


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_2}
ggplot(aes(x =quality , y = alcohol),
       data = df)+
  geom_point(alpha = 1/2)

```

> From the scatter plot above we can see that low quality wine have a less percentage of alcohol in it, and as the quality increases we can see a small correlation between alcohol and the overall quality of the wine, however the highest quality of the wine doesn't have the highest percentage of alcohol .   

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_2.1}
ggplot(aes(x =quality , y = alcohol),
       data = df)+
  geom_point(aes(),stat = 'summary', fun.y = mean,size = 5)

```

>  by the mean alcohol values we can see a clearer picture of the alcohol vs quality and we can say that there is a correlation except for the level 5 quality .

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_3}
ggplot(aes(x = quality, y = pH),
       data = df)+
  geom_point(alpha = 1/5)

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_3.1}
ggplot(aes(x = quality, y = pH),
       data = df)+
  geom_point(aes(alpha = 1/5),stat = 'summary', fun.y = mean,size = 5)

```

> with the mean pH we can see a negative relationship between pH and the quality of the wine except the level 5 quality like the alcohol too.

```{r echo=FALSE,message=FALSE, warning=FALSE,Bivariate_Plots_4}
ggplot(aes(x = quality, y = density),
       data = df)+
  geom_point(alpha = 1/4)

```


> the scatter plot above indecates that high quality wine have less density than low quaity or the average quality wine , however we might take a deeper look on other factors that might effect the quality beside the density .



```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_5}
ggplot(aes(x = quality, y = residual.sugar),
       data = df)+
  geom_point()

```


> both lowest and highest wine quality have alomost the same residual sugar values , with the some outliers in the middle . 


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_6}
ggplot(aes(x = density, y = fixed.acidity),
       data = df)+
  geom_point(alpha = 1/5)+
    geom_smooth()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_cor_test}
cor.test(x=df$density,y=df$fixed.acidity)
```

>  there is a moderate positive relationship between density and fixed acidlity .


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_7}
ggplot(aes(x = volatile.acidity, y = fixed.acidity),
       data = df)+
  geom_point(alpha = 1/2)+
    geom_smooth()

```

>  no strong relationship between fixed acidlity and volatile acidlity .

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_8}
ggplot(aes(x = density, y = residual.sugar),
       data = df)+
  geom_point(alpha = 1/2)+
    geom_smooth()

```


> there is a very small or even insignificant correlation between the residual sugar and density.


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_9}
ggplot(aes(x = density, y = pH),
       data = df)+
  geom_point(alpha = 1/2)+
    geom_smooth()

```

> Comparing pH to density, Most wine samples have a pH between 3.0 to 3.5 , and it doesn't have any effect on the density .


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

> Quality compared to alcohol percentage, low quality wine have a less \
percentage of alcohol in it, and as the quality increases we can see a small\
correlation between alcohol and the overall quality of the wine, however\
the highest quality of the wine doesn't have the highest\
percentage of alcohol .   

>  With The mean pH we can see a negative relationship between pH and\
the quality of the wine.

> The high quality wine have less density than low quaity or the average quality wine.

> conmparing quality to residual sugar, both lowest and highest \
wine quality have alomost the same residual sugar values , with the\
some outliers in the middle . 


### Did you observe any interesting relationships between the other features \

> There is a moderate positive relationship between density and fixed acidlity .

> No strong relationship between fixed acidlity and volatile acidlity .

> There is a very small or even insignificant correlation between the residual sugar and density.

### What was the strongest relationship you found?

> The positive relationship between density and fixed acidlity .

# Multivariate Plots Section



```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
ggplot(aes(x = density, y = fixed.acidity),
       data = df)+
  geom_point(aes(color = quality, alpha = 1/2),stat = 'summary', fun.y = mean,size = 1)

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots2}
ggplot(aes(x = density, y = fixed.acidity),
       data = df)+
  geom_point(aes(color = pH, alpha = 1/2),stat = 'summary', fun.y = mean)
```




```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots3}
ggplot(aes(x = density, y = fixed.acidity),
       data = df)+
  geom_point(aes(color = alcohol, alpha = 1/2),stat = 'summary', fun.y = mean)

```
```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots4}
ggplot(aes(x = density, y = fixed.acidity),
       data = df)+
  geom_point(aes(color = residual.sugar, alpha = 1/2),stat = 'summary', fun.y = mean)

```


> When we added the quality to see it against the fixed acidity vs density , we can not find any corrlation between them and the quality,also alcohol,pH and residual sugar too. 


```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots5}
ggplot(aes(x = quality, y = alcohol),
       data = df)+
  geom_jitter(aes(color = pH, alpha = 1/2),size = 2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots6}
ggplot(aes(x = quality, y = alcohol),
       data = df)+
  geom_jitter(aes(color = density, alpha = 1/2),size = 2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots7}
ggplot(aes(x = quality, y = alcohol),
       data = df)+
  geom_jitter(aes(color = residual.sugar, alpha = 1/2),size = 2)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots8}
ggplot(aes(x = quality, y = alcohol),
       data = df)+
  geom_jitter(aes(color = fixed.acidity, alpha = 1/2),size = 2)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots9}
ggplot(aes(x = quality, y = alcohol),
       data = df)+
  geom_jitter(aes(color = volatile.acidity, alpha = 1/2),size = 2)

```

> > the relationship between alcohol percentage and the quality\
vs (pH,density,residual sugar and both volatile \
and fixed acidity) shows no correlation between all of them.\



```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots10}

m1 <- lm(as.numeric(quality) ~ alcohol, data = df)
m2 <- update(m1, ~ . + pH)
m3 <- update(m2, ~ . + density)
m4 <- update(m3, ~ . + residual.sugar)
m5 <- update(m4, ~ . + fixed.acidity)
m6 <- update(m2, ~ . + volatile.acidity)
mtable(m1,m2,m3,m4,m5,m6,sdigits = 3)

```
> this linear model can only explain 32% of the quality from the  samples in \ 
the red wine dataset with R-squared = 0.323 adding the most features\
that we are interested in (alcohol,pH,density,residual sugar, fixed.acidity and vovolatileacidity) .

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
> there was no features that significantly strength each other in the \
Multivariate Analysis.

> the relationship between alcohol percentage and the quality\
vs (pH,density,residual sugar and both volatile \
and fixed acidity) shows no correlation between all of them.\

> When we added the quality to see it against the fixed acidity vs density ,\
we can not find any corrlation between them and the quality,also alcohol,pH \
and residual sugar too. 

### Were there any interesting or surprising interactions between features?

>The Multivariate Analysis we didn't find any surprising interactions between \
the main features.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

> I have created a linear model to predict the quality of the wine with\
the other features in the data set, and by looking at the results we can \
tell that it only predicts 32% of the quality which is not that significant.


# Final Plots and Summary


### Plot One
```{r echo=FALSE}
ggplot(aes(x = quality, y = alcohol),
       data = df)+
  geom_boxplot(aes(color = quality))+
  ylab("Alcohol Percentage")+
 scale_y_continuous(labels = function(x) paste0(x, "%"))+
 xlab("Quality")+
scale_color_discrete(name = "Quality")+
ggtitle("Boxplot of Alcohol Percentage by Quailty")+
    theme(plot.title = element_text(hjust = 0.5))

```



### Description One
>>The Box plot of  
"Alcohol Percentage by quality
" indicates that there is a correlation between Alcohol \
Percentage and quality and the reason that I decided to \
create a box plot instead of the scatter plot in the univariate\
section is the box plot shows the mean and the range for each\
quality group and it makes it easier to understand the relationship\
between the variables.

### Plot Two

```{r echo=FALSE, Plot_Two}

ggplot(aes(x = density, y = fixed.acidity),
       data = df)+
  geom_point(aes(color = quality))+
  geom_smooth(method = lm)+
 ylab("Fixed Acidity")+
 scale_y_continuous(labels = function(x) paste0(x, " g/dm³"))+
 xlab("Density")+
 scale_x_continuous(labels = function(x) paste0(x, " g/cm³"))+
scale_color_discrete(name = "Quality")+
ggtitle("Scatter Plot of Fixed Acidity by Density with Quality")+
    theme(plot.title = element_text(hjust = 0.5))

```


### Description Two
> this scatter plot shows the correlation between Fixed Acidity and \
the Denisty, and there is a modrete positive relationship between them,\
however when we add the quality to the plot we can see that quality 5 and 6\
are the most common .

### Plot Three
```{r echo=FALSE, Plot_Three}

ggplot(aes(x = quality, y = pH),
       data = df)+
  geom_jitter(aes(color = pH))+
  geom_boxplot(aes(color = pH,alpha = 1/10))+
  ylab("pH")+
  xlab("Quality")+
ggtitle("Jitter Plot of pH by Quality")+
  theme(plot.title = element_text(hjust = 0.5))

```

### Description Three
> This jitter plot is a clearer version of the scatter plot for the pH \
and quality and the added layer of box plots shows the mean and the range\
for the pH for each quality, and we can see a correlation and a negative \
relationship between the pH and the quality.


# Reflection

> The red wine dataset contains information about 1599 samples and 12 variables.
First i examined each individual variable from the dataset to understand the data \
and started asking some questions and there were some trends in the data .
The most important feature of the data is quality of the wine however the vast majority\
of the samples were at level 5 and 6 , so there was a low amount of al lesser or higher \
quality. i
tested lot of combination the bivariant
section and i didn't find a single feature that strongly contributed to the\
quality of the wine except the alcohol we some correlation and the pH we a very small \
correlation. In the multivariant
section i took a deeper look into how features interact with each other and the main\
focus was the quality m but i couldn't find a significant relationship. i
built a linear model that contains the most important feature to see how each one of \
them will explain the quality and how all of them , the results were not very \
informative because it explained only 32% of the quality when we add them all together.\
Limitations include the small number of samples from the high and low quality\
wine and also other factors like The aging of wine which is potentially improved\
the quality, also the county of origin. 
All of these missing information i believe  could explain the quality beside\
the existing variables.

### Sources
Aging of wine   https://en.wikipedia.org/wiki/Aging_of_wine.
adding the % sgin to the plot   https://stackoverflow.com/questions/35967047/appending-symbol-with-y-axis-values-in-ggplot2.
changing the legend name https://stackoverflow.com/questions/33398033/change-ggplot-legend-title.
adjusting to plot title to the cener of the plot https://stackoverflow.com/questions/40675778/center-plot-title-in-ggplot2.

